Binarity's Biases on Transit Survey Occurrence Rates

### 2018-02-24 paper/main.tex
jnw appendix edits

From within this sample of apparently searchable sources, the
observers sum the number of detections $dN_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  dN_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  dN_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  dN_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{dN_{\rm det}}{dr\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this formula
over $r$ after inserting
\begin{equation}
  \Gamma_i(r) = \Lambda(r\p)\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $r\p$ is the
single planet size considered in Section~\ref{sec:model_1}.

@@

Here we present an alternative derivation of
Equation~\ref{eq:general_Gamma_a}.  We also generalize a bit further,
by allowing the occurrence rate density to depend on stellar mass $M$
as well as planet radius $r$.  We begin by writing the apparent rate
density $\Gamma\a$ as (in analogy to
Equation~\ref{eq:occ_rate_simple})
\begin{equation}
    \Gamma\a(r\a,M\a) = \frac{n_{\rm det}(r\a,M\a)}{N_\star(r\a,M\a)}
    \, \frac{1}{p_{\rm tra}(M\a)},
\label{eq:apparent_rate_density}
\end{equation}
where $n_{\rm det}(r\a,M\a)$ is the number of detected planets per
star with radius $r\a$ orbiting stars of apparent mass $M\a$, per unit
$r\a$ and $M\a$.  This collection of planets with the same apparent
properties actually includes planets with a variety of true sizes and
host star masses.  We may write $n_{\rm det}(r\a, M\a)$ as the
convolution of the true occurrence rate density, $\Gamma(r, M)$; and
$\mathcal{N}_\star(r\a, M\a; r, M)$, the number of searchable stars
per unit $(r\a, M\a)$ for which a planet with true properties $(r, M)$
would appear to be a planet with apparent properties $(r\a, M\a)$:
\begin{align}
    n_{\rm det}(r\a, M\a) &=
    \sum_{i=0}^2 n_{\rm det}^i(r\a, M\a) \\
    &=
    \sum_{i=0}^2 \int \mathrm{d}r \int \mathrm{d}M \,
    \mathcal{N}_\star^i(r\a,M\a; r,M)
    \,\Gamma_i(r,M) \, p_{\rm tra}(M),
    \label{eq:n_det}
\end{align}
where $i$ specifies the true type of host star, with 0 for single
stars, 1 for primary stars, and 2 for secondary stars.  We now proceed
to calculate $\mathcal{N}_\star^i(r\a,M\a; r,M)$.

@@

\paragraph{Primaries in binaries}
The number of primaries with apparent parameters $(r\a,M\a)$ given the
true parameters $(r,M)$ is
\begin{equation}
    \mathcal{N}_\star^1(r\a, M\a; r, M)
    = \int 
      \mathrm{d}q\,f(q)\mathcal{N}_{{\rm s}, q}^1(r\a, M\a, q; r, M),
    \label{eq:fancyN_s_1}
\end{equation}
where $f(q)$ is the volume-limited binary mass ratio distribution.
Since we assume that binaries are assigned the mass of the primary,
\begin{equation}
    \mathcal{N}_{{\rm s},q}^1(r\a, M\a, q; r, M)
    \propto
    \hat{\delta}(M\a-M).
\end{equation}
In this case, $\mathcal{N}_{{\rm s},q}^1$ is non-zero only at
$r\a=R\a\sqrt{\delta}$, and the observed depth is
\begin{equation}
    \delta
    = \left[{r\over R(M\a)}\right]^2\, {L(M\a) \over L_{\rm tot}(M\a, q)}
    \equiv \left[{r\over R(M\a)}\right]^2\, 
    \frac{1}{\mathcal{D}_1^2}
    \label{eq:delta_obs_primaries} 
\end{equation}
where
\begin{equation}
    \mathcal{D}_1=\sqrt{L_{\rm tot}(M\a, q) \over L(M\a)}.
\end{equation}
The normalization of $\mathcal{N}_{{\rm s},q}^1$ is given by the
number of binaries that are searchable for a signal $\delta$
and that have mass ratio $q$:
\begin{equation}
    N_\star^0(r\a, M\a)\,
    {n\b\over n\s}\left[L_{\rm tot}(M\a, q) \over L(M\a)\right]^{3/2}.
    \label{eq:primary_normalization}
\end{equation}
where the argument for the number of searchable single stars,
$N_\star^0$, could also be expressed as $(\delta, L(M\a))$.
Applying Equation~\ref{eq:primary_normalization},
\begin{align}
    \mathcal{N}_{{\rm s}, q}^1(r\a, M\a, q; r, M)
    &=N_\star^0(r\a, M\a)
    \, {n\b\over n\s} \, \mathcal{D}_1^3
    \,\hat{\delta} \left(r\a-\frac{r}{\mathcal{D}_1}\right)\hat{\delta}(M\a-M).
    \label{eq:N_s_1}
\end{align}

\paragraph{Secondaries in binaries}
In this case, $M=qM_1=qM\a$, so
\begin{equation}
    \mathcal{N}_{{\rm s},q}^2(r\a, M\a, q; r, M)
    \propto \hat{\delta}\left(M\a-{M\over q}\right).
    \label{eq:fancyN_s_2_prop}
\end{equation}
Again $\mathcal{N}_\star^2$ is non-zero only at
$r\a=R\a\sqrt{\delta}$, but this time
\begin{equation}
      \delta =
      \left[{r\over R(qM\a)}\right]^2 \,
      {L(qM\a) \over L_{\rm tot}(M\a, q)}
      \equiv
      \left[{r\over R(M\a)}\right]^2\,
      \frac{1}{\mathcal{D}_2^2},
\end{equation}
where
\begin{equation}
  \mathcal{D}_2
  =
  \frac{R(qM\a)}{R(M\a)}
  \sqrt{ \frac{L_{\rm tot}(M\a,q)}{L(qM\a)} }.
\end{equation}
The normalization remains the same as the previous case~--~we are
counting the searchable stars at a given observed depth $\delta$, and
the total luminosity of the binary is the same.  Thus,
\begin{equation}
    \mathcal{N}_\star^2(r\a, M\a; r, M)
    =\int \mathrm{d}q\,f(q)\mathcal{N}_{{\rm s}, q}^2(r\a, M\a, q; r, M),
\end{equation}
where
\begin{align}
    \mathcal{N}_{{\rm s},q}^2(r\a, M\a, q; r, M)
    =N_\star^0(r\a, M\a) \, \frac{n\b}{n\s} 
    \, \mathcal{D}_1^3
    \, \hat{\delta} \left(r\a-\frac{r}{\mathcal{D}_2}\right)
    \hat{\delta} \left(M\a-{M\over q}\right).
    \label{eq:mathcal_N_s_2}
\end{align}

One might worry in Equation~\ref{eq:mathcal_N_s_2} that we opt to
write $\mathcal{N}\s^2 \propto \hat{\delta}(M\a - M/q)$, rather than
$\propto \hat{\delta}(M\a q - M)$ or some other delta function with the same
functional dependence, but a different normalization once integrated.
We do this because the delta function in
Equation~\ref{eq:mathcal_N_s_2} is defined with respect to the measure
${\rm d}M\a$, not ${\rm d}M$.  This is because $\mathcal{N}\s^2$ is
defined as a number per $r\a$, per $M\a$.

\paragraph{Number of detected planets}
Marginalizing per Equation~\ref{eq:n_det}, we find
\begin{align}
    \notag
    n_{\rm det}^0(r\a, M\a)
    &=\int\mathrm{d}r\mathrm{d}M\,\mathcal{N}_\star^0(r\a, M\a; r, M)
    \,\Gamma_0(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\,\Gamma_0(r\a, M\a) \, p_{\rm 
        tra}(M\a),
    \label{eq:n_det_0}
\end{align}
and
\begin{align}
    \notag
    n_{\rm det}^1(r\a, M\a)
    &=\int\mathrm{d}r\mathrm{d}M\,\mathcal{N}_\star^1(r\a, M\a; r, M)
    \,\Gamma_1(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \,
    \frac{n\b}{n\s}
    \int \mathrm{d}q \,
         \mathcal{D}_1^3 f(q)\,
         \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a).
    \label{eq:n_det_1}
\end{align}
Finally,
\begin{align}
    n_{\rm det}^2(r\a, M\a)
    \notag
    &=\int\mathrm{d}r\mathrm{d}M\,\mathcal{N}_\star^2(r\a, M\a; r, M)
    \,\Gamma_2(r, M) \, p_{\rm tra}(M)\\
    &=
    N_\star^0(r\a, M\a)\, \frac{n\b}{n\s}
    \int {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\, q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\,
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}.
    \label{eq:n_det_2}
\end{align}

\paragraph{General formula for apparent occurrence rate}
Combining the above results with Equation~\ref{eq:N_searched}, the
apparent rate density,
\begin{align}
    \Gamma\a(r\a,M\a) &= 
    \frac{1}{N_\star(r\a,M\a) p_{\rm tra}(r\a,M\a)} \,
    \sum_i n_{\rm det}^i (r\a,M\a),
\end{align}
evaluates to
\begin{align}
    \notag
    \Gamma\a(r\a,M\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a, M\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\,
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\, q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\,
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}
    \right] \right\}.
\end{align}
We validate this equation in limits where it is possible to write down
the answer (e.g., Equation~\ref{eq:formula_twin_binary}), and also
against a Monte Carlo realization of the twin binary models
(Sections~\ref{sec:model_1} and~\ref{sec:model_2}).

@@

>>>>

From within this sample of apparently searchable sources, the
observers sum the number of detections $dN_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  {\rm d}N_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  {\rm d}N_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  {\rm d}N_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{{\rm d}N_{\rm det}}{{\rm d}r\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this formula
over $r$ after inserting
\begin{equation}
  \Gamma_i(r) = \Lambda(r\p)\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $r\p$ is the
single planet size considered in Section~\ref{sec:model_1}.

@@

Here we present an alternative derivation of
Equation~\ref{eq:general_Gamma_a}.  We also generalize a bit further,
by allowing the occurrence rate density to depend on stellar mass $M$
as well as planet radius $r$.  We begin by writing the apparent rate
density $\Gamma\a$ as (in analogy to
Equation~\ref{eq:occ_rate_simple})
\begin{equation}
    \Gamma\a(r\a,M\a) = \frac{n_{\rm det}(r\a,M\a)}{N_\star(r\a,M\a)}
    \, \frac{1}{p_{\rm tra}(M\a)},
\label{eq:apparent_rate_density}
\end{equation}
where $n_{\rm det}(r\a,M\a)$ is the number of detected planets per
star with radius $r\a$ orbiting stars of apparent mass $M\a$, per unit
$r\a$ and $M\a$.  This collection of planets with the same apparent
properties actually includes planets with a variety of true sizes and
host star masses.  We may write $n_{\rm det}(r\a, M\a)$ as the
convolution of the true occurrence rate density, $\Gamma(r, M)$; and
$\mathcal{N}_\star(r\a, M\a; r, M)$, the number of searchable stars
per unit $(r\a, M\a)$ for which a planet with true properties $(r, M)$
would appear to be a planet with apparent properties $(r\a, M\a)$:
\begin{align}
    n_{\rm det}(r\a, M\a) &=
    \sum_{i=0}^2 n_{\rm det}^i(r\a, M\a) \\
    &=
    \sum_{i=0}^2 \int \mathrm{d}r \!\int\! \mathrm{d}M \,
    \mathcal{N}_\star^i(r\a,M\a; r,M)
    \,\Gamma_i(r,M) \, p_{\rm tra}(M),
    \label{eq:n_det}
\end{align}
where $i$ specifies the true type of host star, with 0 for single
stars, 1 for primary stars, and 2 for secondary stars.  We now proceed
to calculate $\mathcal{N}_\star^i(r\a,M\a; r,M)$.

@@

\paragraph{Primary stars}
For primary stars in binary systems, 
\begin{equation}
    \mathcal{N}_\star^1(r\a, M\a; r, M)
    = \int_0^1 
      \mathrm{d}q\,f(q)\,\mathcal{N}_{{\rm s}, q}^1(r\a, M\a, q; r, M),
    \label{eq:fancyN_s_1}
\end{equation}
where $f(q)$ is the binary mass ratio distribution in a volume-limited sample
of stars.  Since we are assuming that the observer will perceive a binary to be a single
star with the same properties as the primary star,
\begin{equation}
    \mathcal{N}_{{\rm s},q}^1(r\a, M\a, q; r, M)
    \propto
    \hat{\delta}(M\a-M).
\end{equation}
In this case, $\mathcal{N}_{{\rm s},q}^1$ is non-zero only at
$r\a=R\a\sqrt{\delta}$, and the observed transit depth is
\begin{equation}
    \delta
    = \left[{r\over R(M\a)}\right]^2\, {L(M\a) \over L_{\rm tot}(M\a, q)}
    \equiv \left[{r\over R(M\a)}\right]^2\, 
    \frac{1}{\mathcal{D}_1^2}
    \label{eq:delta_obs_primaries} 
\end{equation}
where
\begin{equation}
    L_{\rm tot}(M\a,q) = L(M\a) + L(qM\a),~~\mathcal{D}_1=\sqrt{L_{\rm tot}(M\a, q) \over L(M\a)}.
\end{equation}
The normalization of $\mathcal{N}_{{\rm s},q}^1$ is given by the
number of binaries that are searchable for a signal $\delta$
and that have mass ratio $q$:
\begin{equation}
    N_\star^0(r\a, M\a)\,
    {n\b\over n\s}\left[L_{\rm tot}(M\a, q) \over L(M\a)\right]^{3/2}.
    \label{eq:primary_normalization}
\end{equation}
%where the argument for the number of searchable single stars,
%$N_\star^0$, could also be expressed as $(\delta, L(M\a))$.
Applying Equation~\ref{eq:primary_normalization},
\begin{align}
    \mathcal{N}_{{\rm s}, q}^1(r\a, M\a, q; r, M)
    &=N_\star^0(r\a, M\a)
    \, {n\b\over n\s} \, \mathcal{D}_1^3
    \,\hat{\delta} \left[r\a-\frac{r}{\mathcal{D}_1}\right]\hat{\delta}(M\a-M).
    \label{eq:N_s_1}
\end{align}

\paragraph{Secondary stars}
In this case, $M=qM_1=qM\a$, so
\begin{equation}
    \mathcal{N}_{{\rm s},q}^2(r\a, M\a, q; r, M)
    \propto \hat{\delta}\left(M\a-{M\over q}\right).
    \label{eq:fancyN_s_2_prop}
\end{equation}
Again $\mathcal{N}_\star^2$ is non-zero only at
$r\a=R\a\sqrt{\delta}$, but this time
\begin{equation}
      \delta =
      \left[{r\over R(qM\a)}\right]^2 \,
      {L(qM\a) \over L_{\rm tot}(M\a, q)}
      \equiv
      \left[{r\over R(M\a)}\right]^2\,
      \frac{1}{\mathcal{D}_2^2},
\end{equation}
where
\begin{equation}
  \mathcal{D}_2
  =
  \frac{R(qM\a)}{R(M\a)}
  \sqrt{ \frac{L_{\rm tot}(M\a,q)}{L(qM\a)} }.
\end{equation}
The normalization remains the same as the previous case~--~we are
counting the searchable stars at a given observed depth $\delta$, and
the total luminosity of the binary is the same.  Thus,
\begin{equation}
    \mathcal{N}_\star^2(r\a, M\a; r, M)
    =\int_0^1 \mathrm{d}q~f(q)\,\mathcal{N}_{{\rm s}, q}^2(r\a, M\a, q; r, M),
\end{equation}
where
\begin{align}
    \mathcal{N}_{{\rm s},q}^2(r\a, M\a, q; r, M)
    =N_\star^0(r\a, M\a) \, \frac{n\b}{n\s} 
    \, \mathcal{D}_1^3
    \, \hat{\delta} \left(r\a-\frac{r}{\mathcal{D}_2}\right)
    \hat{\delta} \left(M\a-{M\over q}\right).
    \label{eq:mathcal_N_s_2}
\end{align}

\paragraph{Number of detected planets}
Inserting the results into Equation~\ref{eq:n_det}, we find
\begin{align}
    \notag
    n_{\rm det}^0(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^0(r\a, M\a; r, M)
    \, \Gamma_0(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, \Gamma_0(r\a, M\a) \, p_{\rm 
        tra}(M\a),\\
    \notag
    n_{\rm det}^1(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^1(r\a, M\a; r, M)
    \, \Gamma_1(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \,
    \frac{n\b}{n\s}
    \int \mathrm{d}q \,
         \mathcal{D}_1^3 f(q)\cdot
         \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a),\\
    n_{\rm det}^2(r\a, M\a)
    \notag
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^2(r\a, M\a; r, M)
    \, \Gamma_2(r, M) \, p_{\rm tra}(M)\\
    &=
    N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \, \frac{n\b}{n\s}
    \int {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\cdot
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}.
    \label{eq:n_det_2}
\end{align}

Finally, combining these results, the
apparent occurrence rate density
\begin{align}
    \Gamma\a(r\a,M\a) &= 
    \frac{1}{N_\star(r\a,M\a) p_{\rm tra}(r\a,M\a)} \,
    \sum_{i=0}^2 n_{\rm det}^i (r\a,M\a)
\end{align}
becomes
\begin{align}
    \notag
    \Gamma\a(r\a,M\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a, M\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\cdot
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}
    \right] \right\}.
\end{align}
%We validate this equation in limits where it is possible to write down
%the answer (e.g., Equation~\ref{eq:formula_twin_binary}), and also
%against a Monte Carlo realization of the twin binary models
%(Sections~\ref{sec:model_1} and~\ref{sec:model_2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
